---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# DoEAna

<!-- badges: start -->
<img src  = "vignettes/badge.png" align = "right" height = "135" />
<!-- badges: end -->

The goal of DoEAna is to visualize the results of research regarding 
space-filling experimental designs. The package has a prepared data set
"DoEAna.rda" with designs for comparison. In the dataset, you can see how
designs were generated on two different geometries: hypercube and simplex. They
also differ by the distance measure used: Manhattan, Chebyshev, and Euclidean. 
Finally, they differ by the number of design points: 10, 20, and 30. The user 
can use this package to compare how these different experiment features affect
a designs performance. Although the main purpose of this package is to display
specific results of a research project, these functions can be used by any
individual that formats their data in the same way as DoEAna.rda in the package
data folder.




## Installation

You can install the development version of DoEAna from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("DianaF65/DoEAna")
```

## Examples

This is a basic example which shows you how to use this package:

First, select optimal space-filling designs. You can view the design matrix and plot them. 

```{r example}
library(DoEAna)

# select designs on the hypercube or simplex
my_design_hyper <- select_design(d_data = sf_designs, geometry = "hypercube", distance = "chebyshev", n = 20)
my_design_sim <- select_design(d_data = sf_designs, geometry = "simplex", distance = "chebyshev", n = 20)

# show the design matrices
print_design(my_design_hyper)
print_design(my_design_sim)

# plot the designs
par(mfrow = c(1, 2))
plot_design(my_design_hyper)
plot_design(my_design_sim)

```

Investigate their relative efficiency. 


```{r}
# compare across different distance measures with relative efficiency on the 
# hypercube
my_designM_H <- select_design(d_data = sf_designs, geometry = "hypercube",
                            distance = "manhattan", n = 20)
my_designE_H <- select_design(d_data = sf_designs, geometry = "hypercube",
                            distance = "euclidean", n = 20)
my_designC_H <- select_design(d_data = sf_designs, geometry = "hypercube",
                            distance = "chebyshev", n = 20)

relative_efficiency(m_row = my_designM_H, e_row = my_designE_H, 
                    c_row = my_designC_H)
```

```{r cars}
# compare across different distance measures with relative efficiency on the 
# simplex
my_designM_S <- select_design(d_data = sf_designs, geometry = "simplex",
                            distance = "manhattan", n = 20)
my_designE_S <- select_design(d_data = sf_designs, geometry = "simplex",
                            distance = "euclidean", n = 20)
my_designC_S <- select_design(d_data = sf_designs, geometry = "simplex",
                            distance = "chebyshev", n = 20)

relative_efficiency(m_row = my_designM_S, e_row = my_designE_S, 
                    c_row = my_designC_S)
```

Or view an FDS plot

```{r}
# Compare relative prediction variance across designs using FDS plots
fds_plot(design1 = my_designC_H, design2 = my_designM_H, 
         title = "Chebychev vs Manahattan Distance for n=20")
```

